---
title: "Explanatory Model Analysis"
author: "R. J. Serrano"
date: "2024-06-20"
output: 
     html_document:
          df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE,
                      message = FALSE, dpi = 180, 
                      fig.width = 8, fig.height = 5)

suppressMessages(library(tidyverse))
library(scales)
library(DALEX)
library(DataExplorer)
theme_set(theme_bw())
```

## Chapter 21 - FIFA 19 Dataset

The objective is to explore the FIFA 19 dataset included in the `DALEX` package [original dataset](https://www.kaggle.com/datasets/javagarm/fifa-19-complete-player-dataset)

```{r fifa19}
load('./data/fifa19small.rda')

fifa <- fifa19small |> 
     rename(Reputation = 9) |> 
     remove_rownames() |> 
     column_to_rownames(var = "Name")

fifa
```

First look at datatset (skim)
```{r skim}
skimr::skim(fifa)
```

Number of observations (rows): 16,924

Number of variables (columns): 44 (index = player's name)

Number of numeric/continuous variables: 41

Number of categorical variables: 3 (`Club`, `Position`, `Preferred.Foot`)

Number of missing values: 0

Target variable (response): `value_eur`

Histogram of player's value (`Value.EUR`)
```{r histogram-value-eur}
fifa |> 
     ggplot(aes(Value.EUR)) + 
     geom_histogram(fill = 'steelblue')
```

Histogram of log-transformed player's value (`Value.EUR`)
```{r log-trans-histogram-value-eur}
fifa |> 
     ggplot(aes(log10(Value.EUR + 0.1))) + 
     geom_histogram(fill = 'steelblue')
```

Histograms for selected variables (figure 21.2)
```{r selected-vars-histograms}
fifa |> 
     select(Age, BallControl, Dribbling, Reactions) |> 
     plot_histogram(nrow = 2L, ncol = 2L)
```

1. Most players' age fall between 20 and 40 years old.
2. Smaller modes in `BallControl` and `Dribbling` are due to goalkeepers.

Scatterplots for selected variables (figure 21.3)
```{r selected-vars-scatterplots}
fifa19small4long <- fifa19small |> 
     select(Value.EUR, Age, BallControl, Dribbling, Reactions) |> 
     gather(variable, value, -Value.EUR)

fifa19small4long |> 
     ggplot(aes(value, Value.EUR)) + 
     geom_point() + 
     geom_smooth(linewidth = 2, se = FALSE, color = 'steelblue') + 
     theme_drwhy() + 
     facet_wrap(~variable, ncol = 2, scales = "free") + 
     scale_y_continuous("Value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
     scale_x_continuous("") + 
     ggtitle("Scatterplots for players' characteristics","") + 
     theme_ema()
```

Scatterplot matrix for selected variables
```{r selected-vars-scatterplot-matrix}
library(GGally)

fifa19small |> 
     select(Age, BallControl, Dribbling, Reactions) |> 
     ggpairs(
          diag = list(continuous = "barDiag"), 
          lower = list(continuous = wrap("smooth", method = "loess", colour = 'steelblue'))
          ) + 
     theme_drwhy() + 
     ggtitle("Scatterplot matrix for players' characteristics","") + 
     theme_ema()
```

Correlation funnel
```{r corr-funnel}
library(correlationfunnel)

# binarized features
binarized_tbl <- fifa %>% 
     select(-Club, -Overall) |> 
     binarize(n_bins = 5, thresh_infreq = 0.01, name_infreq = 'Other', one_hot = TRUE)

# correlate to target ('Value.EUR__3100000_Inf')
corr_tbl <- binarized_tbl %>% 
     correlate(Value.EUR__3100000_Inf)

# plot correlation
corr_tbl %>% 
     plot_correlation_funnel(interactive = TRUE)
```
